RProtoBuf: Efficient Cross-Language Data Serialization in R

### 2014-01-03 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Further improve the introduction.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@697 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Programming languages such as Java, Ruby, Python, and R include
built-in serialization support, but these formats are tied to the
specific programming language in use and thus lock the user into a
single environment.
%
% do not facilitate
% TODO(ms): and they often don't support versioning among other faults.
CSV files can be read and written by many applications and so are
often used for exporting tabular data.  However, CSV files have a
number of disadvantages, such as a limitation of exporting only
tabular datasets, lack of type-safety, inefficient text representation
and parsing, and ambiguities in the format involving special
characters.  JSON is another widely-supported format used mostly on
the web that removes many of these disadvantages, but it too suffers
from being too slow to parse and also does not provide strong typing
between integers and floating point.  Because the schema information
is not kept separately, multiple JSON messages of the same
type needlessly duplicate the field names with each message.
%
%
%
A number of binary formats based on JSON have been proposed that
reduce the parsing cost and improve the efficiency.  MessagePack
\citep{msgpackR} and BSON \citep{rmongodb} both have R interfaces, but
these formats lack a separate schema for the serialized data and thus
still duplicate field names with each message sent over the network or
stored in a file.  Such formats also lack support for versioning when
data storage needs evolve over time.

@@

Once the data serialization needs get complex enough, application
developers typically benefit from the use of an \emph{interface
description language}, or \emph{IDL}.  IDLs like Google's Protocol
Buffers and Apache Thrift provide a compact well-documented schema for
cross-langauge data structures as well efficient binary interchange
formats.  The schema can be used to generate model classes for
statically typed programming languages such as C++ and Java, or can be
used with reflection for dynamically typed programming languages.
Since the schema is provided separately from the encoded data, the
data can be efficiently encoded to minimize storage costs of the
stored data when compared with simple ``schema-less'' binary
interchange formats like BSON.

>>>>

Given these requirements, how do we safely share intermediate results
between different applications, possibly written in different
languages, and possibly running on different computers?  Programming
languages such as R, Java, Julia, and Python include built-in
serialization support, but these formats are tied to the specific
programming language in use and thus lock the user into a single
environment.  CSV files can be read and written by many applications
and so are often used for exporting tabular data.  However, CSV files
have a number of disadvantages, such as a limitation of exporting only
tabular datasets, lack of type-safety, inefficient text representation
and parsing, and ambiguities in the format involving special
characters.  JSON is another widely-supported format used mostly on
the web that removes many of these disadvantages, but it too suffers
from being too slow to parse and also does not provide strong typing
between integers and floating point.  Because the schema information
is not kept separately, multiple JSON messages of the same type
needlessly duplicate the field names with each message.
%
%
%
A number of binary formats based on JSON have been proposed that
reduce the parsing cost and improve the efficiency.  MessagePack
\citep{msgpackR} and BSON \citep{rmongodb} both have R interfaces, but
these formats lack a separate schema for the serialized data and thus
still duplicate field names with each message sent over the network or
stored in a file.  Such formats also lack support for versioning when
data storage needs evolve over time.

Once the data serialization needs of an application become complex
enough, developers typically benefit from the use of an
\emph{interface description language}, or \emph{IDL}.  IDLs like
Google's Protocol Buffers, Apache Thrift, and Apache Avro provide a compact
well-documented schema for cross-langauge data structures and
efficient binary interchange formats.  The schema can be used to
generate model classes for statically typed programming languages such
as C++ and Java, or can be used with reflection for dynamically typed
programming languages.  Since the schema is provided separately from
the encoded data, the data can be efficiently encoded to minimize
storage costs of the stored data when compared with simple
``schema-less'' binary interchange formats.

@@
